---
title: "SDS 201 Data Analysis Project"
author: "Musayeroh Bah and Lily Mitchell"
format: 
    html:
      embed-resources: true 
code-fold: false
execute: 
  echo: true
knitr: 
  opts_chunk: 
    message: false
    warning: false 
editor: source 
bibliography: proj_refs.bib
---

```{r}
#| eval: true
#| echo: false
#| message: false

library(tidyverse)
library(googlesheets4)
library(ggplot2)
library(dplyr)
library(janitor)
library(moderndive)
library(knitr)
library(infer)
```

![](https://img10.hotstar.com/image/upload/f_auto/sources/r1/cms/prod/8995/1368995-i-b1fa2d292a86){width="900"}

# Hollywood vs. The Bechdel Test

Disparities in popular entertainment have a long history rooted in gender. Gender bias disproportionately affects the number of films the industry produces that include women in lead roles or substantive ones. Growing media consumption for all ages, highlights the importance of representation in landscapes mapped by media and their broader impact. Attesting to this, in 2023, Geena Davis, spoke of media's influence where “Hollywood creates our cultural narrative – its biases trickle down to the rest of the world” and the responsibility of those both creating and funding content to challenge these systemic issues [@davis2023hollywood].

Our research project takes a deeper look at these inequities and analyzes 1794 popular movies produced over the span of 43 years. In order to assess and quantify the evidence of gender bias and its monetary consequences, we utilize the measures best captured by the Bechdel-Wallace Test. Commonly abbreviated as the Bechdel Test, it was designed by Alison Bechdel to measure representation of women in film and other fictional media. For a movie to pass this test it must satisfy all three requirements of featuring two female characters, who have a conversation, about a topic other than men. Using the [@fivethirtyeight_bechdel] data set we explore the industry's inequities and provide insights into the financial dynamics of representation in film.\

It may not come as a surprise that our analysis reveals the dramatic extent to which budget differences are tied to Bechdel Test results. Our findings suggest that movies failing to pass the Bechdel Test are often made with bigger budgets and consequently lack meaningful representation of women and strong female characters. This points to a potential systemic bias in film and media production and distribution of resources that impacts consumers of content and larger cultural attitudes towards women.

## Exploratory Data Analysis

Our data, as introduced above, includes information about popular movies and whether or not they pass the Bechdel test. This data set holds a sample of 1,794 films released between the years of 1970 and 2013 [@hickey2014dollars].

In wrangling the data, we filtered the dataset so that only the variables most relevant to our analysis, “title”, “binary”, and “budget” were kept. Additionally, instead of using the budget column, we filtered for the `budget_2013` variable as it had all of the budgets adjusted for inflation, in 2013 US dollars, making each movie's budget comparable against each other. The title variable reveals the name of the movie and the binary variable column represents, for each movie, pre-identified values of PASS/FAIL calculated with the Bechdel Test.\

Given the gender bias evident from the Bechdel Test results, our data analysis considers the possible relationship between the numeric variable of budget and the categorical variable of Pass/Fail to underrepresentation. To investigate this, we pose the question: "Do movies that pass the Bechdel Test have a higher budget?".

The numerical variable Budget serves as the outcome variable in our analysis while the categorical variable, Binary, that denotes Passing/Failing the Bechdel Test the explanatory variable of the analysis. To test the strength and magnitude of this relationship, we used a regression model with a permutation test.

### Data Wrangling

@tbl-movies_raw displays the raw data we imported from @fivethirtyeight_bechdel. @tbl-selected_moviedata displays the variables that we included for our analysis, and in @tbl-movies_clean_budget, the Budget variable has been wrangled and updated to reflect the 2013 inflation adjusted budget.\\

#### Imported Data

```{r}
#| label: tbl-movies_clean_budget
#| tbl-cap: "Cleaned Dataset that includes all budgets adjusted for inflation, in 2013 USD"
#| eval: true
#| echo: false
#| message: false
movies_raw <- read_sheet("https://docs.google.com/spreadsheets/d/12xXTpjqwY35HV5WcAZtIGxnLc8xmBD0hQPVZKR5njcY/edit?gid=1554984819#gid=1554984819")
movies <- movies_raw |>
  select(c("title", "binary", "budget_2013"))

movies_clean <- movies_raw |>
  select(c("title", "binary", "budget_2013")) |>
  mutate(budget_2013 = as.numeric(budget_2013))
  
movies_show <- head(movies_clean, n=3)
 kable(movies_show)
```


### Summary of Data

In @tbl-movies_summary we see the breakdown of average budgets for movies that either pass or fail the Bechdel test. The standard deviation is given along with the measure of central tendency and the Number column reflects the number of movies in each category contained in the dataset. \ 

```{r}
#| label: tbl-movies_summary
#| tbl-cap: "Summary of Movies Data "
#| eval: true
#| echo: false
#| message: false
plotting <- movies_clean |>
  group_by(binary) |>
  summarize(avg_budget = mean(budget_2013),
           Number = n(),
           Standard_Deviation = sd(budget_2013))
kable(plotting)
```

@fig-movies_histogram visualizes the quantitative difference between the average budgets. This original finding prompted our hypothesis test to further explore a possible link between movie budgets and passing or failing the Bechdel test.\ 

In @fig-movies_boxplot we can see the of this difference in average budget more precisely with a box plot that captures the distribution of budgets across movies with outliers and percentiles.\ 

```{r}
#| label: fig-movies_histogram
#| fig-cap: "Average Budget Differences"
#| eval: true
#| echo: false
#| message: false
ggplot(plotting, aes(x=binary, y=avg_budget, fill = binary)) +
  geom_bar(stat = "identity") +
  scale_y_continuous(labels = scales::comma)+
  labs(x = "PASS / FAIL the Bechdel Test ",
       y= "Average Budget",
       title = "Average Budget Differences based on Bechdel Test Result"
       )
```

```{r}
#| label: fig-movies_boxplot
#| fig-cap: "Boxplot showing the difference between average movie budgets more accuratly capturing outliers"
#| eval: true
#| echo: false
#| message: false
ggplot(movies_clean, aes(x=binary, y=budget_2013, fill = binary)) +
  geom_boxplot() +
  scale_y_continuous(labels = scales::comma)+
  labs(x = "PASS / FAIL",
       y= "Movie Budgets Reflecting USD of 2013",
       title = "Difference Between Average Movie Budgets Capturing Outliers"
       )
```

```{r}
#| label: fig-budget_Histogram_distribution
#| fig-cap: "Distribution of Difference in Budgets that PASS vs FAIL the Bechdel Test"
#| eval: true
#| echo: false
movies_clean |>
  ggplot(aes(x=budget_2013, fill = binary))+
  geom_histogram()+
  theme_minimal() +
  scale_y_continuous(labels = scales::comma)+
  labs(x = "Budget",
       y= "PASS vs FAIL",
       title = " Distribution of Difference in Budgets that PASS vs FAIL the Bechdel Test"
       )
```


## Results

To exactly quantify the relationship between the average budget of a movie and whether it passes or fails the Bechdel test, we fit a categorical regression model. In the model portrayed by @tbl-regression_table, the outcome variable is the budget of a movie and the explanatory variable is the movie passing or failing the aforementioned test. The predictions created from the model predictions are visualized in @fig-Categorical_vizualization.

### Hypothesis Testing

#### Categorical regression model

```{r}
#| label: tbl-regression_table
#| tbl-cap: "Regression Table"
#| eval: true
#| echo: false
#| message: false
bechdel_model <- lm(budget_2013 ~ binary, data = movies_clean)

bechdel_model_clean <- get_regression_table(bechdel_model)

kable(bechdel_model_clean)
```

#### Movie Budgets vs. Bechdel Test Binary

```{r}
#| label: fig-Categorical_vizualization
#| fig-cap: "Catagorical Visualization of difference between average movie budgets "
#| eval: true
#| echo: false
#| message: false
ggplot(bechdel_model, aes(x = binary, y = budget_2013, color = binary)) +
  geom_jitter(width = 0.25) +
  geom_categorical_model() +
  theme_minimal() +
  scale_y_continuous(labels = scales::comma)+
  labs(title = "Movie Budgets vs. Bechdel Test Binary",
       x = "Binary - PASS/FAIL",
       y = "Budget (in 2013 USD)")
```

The @tbl-regression_table fits our hypothesis test to the data and gives us a table of values. The probability of observing a test statistic at least as extreme as the one observed from our sample, assuming the null hypothesis is true, is 0. Our alpha level is 0.05, and our p-value is 0, and because p-value \< alpha value, we can reject the null hypothesis and conclude that there is a statistical significance between the average budget and a movie passing or failing the bechdel test. 

One can see in @fig-Categorical_vizualization the Catagorical Visualization of difference between average movie budgets.\ 

#### Regression Equation: 

The regression equation created/described by the Table is:
\ 
$$
\widehat{Budget} = \widehat{\beta_0} + \widehat{\beta_1} \cdot 1_{Pass}(Binary)
$$
\ 
$$
\widehat{Budget} = 62911555 + (-16637388 * 1_{Pass}(Binary))
$$ 
\
The intercept, ($\widehat{\beta_0} = 62911555$) represents the average budget of a movie when it passes the bechdel test, and $\widehat{\beta_1}$, represents the decrease in the average budget when a movie fails the bechdel test, which is -16637388.

Although this might be due to chance, a permutation test on the model’s slope coefficient shows that there is not strong enough evidence to conclude that the differences between the average budgets are produced simply due to variability between the sample we used and others (t = 16637388, p = 0).

### Discussion

#### Conclusions

Through summarizing the data, we have gathered that, after adjusting for inflation in 2013 USD, the average budget of movies that fail the Bechdel test tends to be higher than the average budget of movies that pass it. This suggests that there's a possible link between how much money is ensured to movies that largely suffer from a lack of gender representation upon release and those that don't. 

-   talk about more insights with a more specific look at how other variables may be affected; do certain genres suffer from failing the test more than others? - talk about how it could be for many reasons that we, the researchers are still unsure about, like it could be that writers make deliberate decisions to give importance to main male leads rather than female leads, sexism within writing scripts, sexism from writers and/or executives that help make decisions on budgets. But it is a problem if films without big budgets behind them have actual 3-dimensional characters. Overall

#### Limitations

There are several limitations to the dataset and our findings that lead us to say that more research on the topic is needed to make any definitive conclusions. For starters, this dataset only uses movies from 1990 to 2013. Also, this dataset doesn't distinguish international movies such as Monty Python and The Holy Grail (1975) and Ponyo (2008) to US films in terms of budget. It does not give the best picture to analyze these movies in USD compared to their own country’s currency due to exchange rates and what’s considered a ‘high’ and ‘lower’ budget in their place of origin. Therefore, the results of this study can't be generalized entirely to other film industries. It also isn't a good look at the US film industry itself before 1990. 

(mmm keep this part or no?)
These results also don’t model how well the movie passes the test or provides insight to that; some movies that pass may consist of compelling and plot-relevant female characters that have conversations throughout the movie or simply have a single moment of them talking in an hour runtime about something irrelevant to the plot. The test is simple, and while a movie not passing it says a lot for the representation of women in that film (whether it exists or not), a movie passing doesn’t really say a lot about the quality of that representation. More qualitative research should be undertaken to gain a deeper insight.

Additionally, it is important to note that this study should not take away from the fact that there are many factors that go into creating a budget, and we cannot claim that dimensionality of female characters is the sole or even main reason that a movie obtains the budget that it does. Other considerations, like costuming and set design, are huge considerations for any movie's budget. 

### Further Questions
With the aforementioned limitations, a good place to start with further analysis would be investigating if movies produced outside of the US fare better in terms of budget (in their country-of-origin’s economy) and measuring up against the test. 

On the last point, it should not understated the importance of doing further quantitative analysis into how other factors link to a movie’s ability to pass or fail this test. For example, should a movie pass the Bechdel test, does it bring in more or less domestic and/or international revenue than if it failed? That could provide insight into how well certain movies are received by the public both within the US and outside of it.

Figure @fig-movies_domgross shows a preliminary exploratory analysis of how domestic gross varies between films that pass/fail the Bechdel test. Starting off, this figure shows that movies that pass the Bechdel test with a higher budget also earn more domestically than other movies, while those that get higher budgets in the 'FAIL' category seem to earn revenue at a lesser rate. A deeper analysis would 


```{r}
#| label: fig-movies_domgross
#| fig-cap: "Difference Between Movie Budget and Gross Domestic Revinue"
#| eval: true
#| echo: false
#| message: false
movies_raw |> 
  filter(!is.na(domgross_2013)) |>
  select(binary, budget_2013, domgross_2013) |>
  mutate(budget_2013 = as.numeric(budget_2013),
         domgross_2013 = as.numeric(domgross_2013))|>
  
ggplot(aes(x=domgross_2013, y= budget_2013, color = binary))+
  geom_point() +
  geom_smooth(method ="lm", se = FALSE) +
  facet_wrap(vars(binary), ncol = 2)+
  labs(x = "Gross Domestic 2013",
       y= "Budget 2013",
       title = "Difference Between Movie Budget and Gross Domestic Revinue"
       )
```

#### Bibliography


### Appendix 

### Null Hypothesis: \ 
There is no difference in the average budget of movies that pass or fail the Bechdel test. 
$$
\mu_{Pass} - \mu_{Fail} = 0
$$



### Alternative Hypothesis:\
There is a difference in the average budget of movies that pass or fail the Bechdel test.
$$
\mu_{Pass} - \mu_{Fail} \ne 0
$$


### Permutation distribution
```{r}
#| eval: true
#| echo: false
#| message: false
library(infer)
set.seed(123)

permutation_distribution <- movies_clean |>
specify(budget_2013 ~ binary) |>
hypothesise(null = "independence") |>
generate(reps = 1000, type="permute") |>
calculate(stat = "diff in means",
          order = c("PASS", "FAIL"))
visualize(permutation_distribution)
```

### P-value
```{r}
#| eval: true
#| echo: false
#| message: false
permutation_distribution |>
  get_p_value(obs_stat = 16637388,
              direction = "both")
```

### Shaded P-Value
```{r}
#| eval: true
#| echo: false
#| message: false
visualize(permutation_distribution) + shade_p_value(obs_stat = 16637388, direction = "both")
```

```{r}
#| label: fig-summary_stat
#| fig-cap: "Calculating the Summary Statistic for Hypothesis Test "
#| eval: false
#| echo: false
#| message: false
62911555 - 46274167
```
