---
title: "Data_Analysis Project"
author: "Musa Bah and Lily Mitchell"
bibliography: references.bib
csl: apa-numeric-superscript.csl
format:
  html:
    embed-resources: true
    df-print: paged
    tbl-cap-location: bottom
execute: 
  echo: false
  warning: false
  message: false
knitr:
  opts_chunk: 
    dev: svg
    cache: true
---

```{r}
#| label: libraries
library(tidyverse)
library(googlesheets4)
library(janitor)
library(moderndive)
library(knitr)
library(gridExtra)
library(GGally)
```

Core Requirements:

Your analysis must also include at least one hypothesis test that
supports an inference about how your outcome variable is related to your
explanatory variable(s) in the population.

● For example, you may include a hypothesis test for a difference in
means, a hypothesis test for a difference in proportions, or a
hypothesis test on the coefficients of a regression model.

● You may use either randomization-based hypothesis tests (e.g., a
permutation test) or a parametric hypothesis tests (e.g., a t-test).

Your data analysis project will address a question like “How is
{Variable X} related to {Variable Y}?”.

● This means that your data set must measure several variables

● You must choose one variable to serve as your outcome variable

● You must choose another variable (or variables) to serve as
explanatory variable(s).

Your report should also cite at least one outside source of information
(in addition to citing the source of your data), though that source does
not necessarily have to be a peer-reviewed “academic” article.

# Introduction

Introduction: an overview of your project. In a few paragraphs, you
should explain clearly and precisely what your research question is, why
it is interesting, and what contribution you have made towards answering
that question. You should give an overview of the specifics of your
model, but not the full details. Most readers never make it past the
introduction, so this is your chance to hook the reader, and is in many
ways the most important part of the paper!

Introduces the topic of your analysis (including your research
question(s) and the goals of your data analysis)

The Bechdel Test is used to measures representaiton of women in film and
other fictional media. For a movie to pass the test it must feature two
female characters who have converstations about a topic other than men.

● Explains the nature of the data you are analyzing in sufficient detail

● Explains what analyses are performed and what can be learned from them

● Ends with overall conclusions related to your research question that
are supported by your data analysis.

What research question do you plan to address using these variables?

-   Do movies that pass the Bechdel Test have a higher budget?

Describe your hypothesis about how these variables are related to one
another.

-   The movies with lower budgets fail the bechdel test.

Which variable will be the outcome variable in your analysis? Is this a
categorical variable or a numeric variable? If categorical, describe the
levels of this categorical variable

-   Budget

Which variable will be the explanatory variable in your analysis? Is
this a categorical variable or a numeric variable?

-   Pass or Fail Bechdel Test

What method will you use to test your hypothesis (two-sample t-test,
regression model with a permutation test, etc.)

-   Regression model with a permutation test

## Data

Data: a brief description of your data set. This section should
introduce the reader to the data by answering questions such as: what
variables are included? Where did they come from? What are units of
measurement? What is the population that was sampled? How was the sample
collected? This section should also acquaint the reader with your data
set through the use of univariate and multivariate visualizations, and
univariate and multivariate summaries.

This section should aldo describe any important “data wrangling” (e.g.,
filtering missing values, recoding variables) you did throughout the
course of your analysis. The key word here is “important”. Identifying
what is important can be difficult at first. Try asking yourself these
guiding questions to help you decide:

◦ Could my conclusions be different I had/had not done this part of my
data wrangling?

◦ Is it reasonable that a reader could view my work differently if they
knew, or didn't know about this?" For example

◦ Could my conclusions be different I had/had not removed missing
observations? Yes, definitely, so it should be described.

◦ Could my conclusions be different if I described how the missing
observations were removed? Probably not (using dplyr::filter has the
same overall consequences as deleting rows in a spreadsheet, so it’s not
worth mentioning)

◦ Is it reasonable that a reader could view my work differently if they
knew I chose to analyze a variable related to revenue when I could have
analyzed a variables related to profit? Yes, so it’s worth mentioning I
chose this specific variable and not another.

◦ Is it reasonable that a reader could view my work differently if they
knew I renamed the variables from all capitals to all lower case?
Definitely not

```{r}
movies <- read_sheet("https://docs.google.com/spreadsheets/d/12xXTpjqwY35HV5WcAZtIGxnLc8xmBD0hQPVZKR5njcY/edit?gid=1554984819#gid=1554984819")
movies
```

```{r}
movies |>
  select(title, binary, budget)
```

## Results

Results: A hypothesis test set up to address your research questions(s).

You should interpret the results in the context of the data explain
their relevance. If you are using a regression model in your analysis,
each of it’s coefficients should be clearly explained.

You should include negative results, but be careful about how you
interpret them. For example, you may want to say something along the
lines of: “we found no evidence that explanatory variable x is
associated with response variable y,” or “explanatory variable x did not
provide any additional explanatory power above what was already conveyed
by explanatory variable z”. On other hand, you probably shouldn’t claim:
“there is no relationship between x and y”.

## Discussion

Discussion/Conclusion: a summary of your findings and a discussion of
their limitations. Be sure to remind the reader of the question that you
originally set out to answer, and summarize your answer to the question.
Your discussion should also protect yourself against misinterpretation
by being clear about what is not implied by your research. Finally, you
should also discuss the limitations of your analysis and/or data, and
how it could be augmented or improved with future research. This “future
questions” portion of the discussion should include a brief exploratory
analysis showing how your project could be extended to include
additional explanatory variables in the future.

### Conclusions

### Limitations

### Further Questions

#### Appendix

Appendix (Potentially Optional): If you have any supplemental analyses
that would be of interest (e.g., examining the assumptions of your
two-sample t-test), you may wish to place those analyses in an appendix
at the end of your report.

#### Bibliography

(e.g., APA or MLA format)

Write your technical report as an Rmarkdown or Quarto document, the PDF
and HTML file you submit should not display any R code, warnings, or
messages. Your technical report should also be well formatted and well
organized (e.g., using properly formatted headers to divide sections,
using LaTeX markup for mathematical symbols where needed, tables should
have bolded column and row labels, etc.).
