---
title: "SDS 201 Data Analysis Project"
author: "Musayeroh Bah and Lily Mitchell"
format: 
    html:
      embed-resources: false
code-fold: false
execute: 
  echo: true 
knitr: 
  opts_chunk: 
    message: false
    warning: false 
editor: source 
bibliography: proj_refs.bib
---


```{r}
#| eval: true
#| echo: false
#| message: false

library(tidyverse)
library(googlesheets4)
library(ggplot2)
library(dplyr)
library(janitor)
library(moderndive)
library(knitr)
```

![BechdelTest](https://img10.hotstar.com/image/upload/f_auto/sources/r1/cms/prod/8995/1368995-i-b1fa2d292a86){width="900"}


# Hollywood vs. The Bechdel Test

Disparities in popular entertainment have a long history rooted in gender. Gender bias disproportionately affects the number of films the industry produces that include women in lead roles or with roles as substantive characters. Growing media consumption for all ages, highlights the importance of representation in landscapes mapped by these media and their broader impact. In 2023, Geena Davis, attesting to this, spoke of the influence where “Hollywood creates our cultural narrative – its biases trickle down to the rest of  the world” and the responsibility of those both creating and funding content to challenge these systemic issues @davis2023hollywood. 

Our research project takes a deeper look at these inequities and analyzes 1794 popular movies produced over the span of 43 years. In order to assess and quantify the evidence of gender bias and its monetary relationship we utilize the measures best captured by the Bechdel-Wallace Test. Commonly abbreviated as Bechdel Test, it was designed by Alison Bechdel to measure representation of women in film and other fictional media. For a movie to pass this test it must satisfy all three requirements of: featuring two female characters, who have a conversation, about a topic other than men. Using the @fivethirtyeight_bechdel data set we explore industry's inequities and provide insights into the financial dynamics of representation in film.\

It may not come as a surprise that our analysis reveals the dramatic extent to which budget differences are tied to Bechdel Test results. Our findings suggest that movies failing to pass the Bechdel Test are often made with smaller budgets and therefore lack meaningful representation of women and strong female characters. This points to a potential systemic bias in film and media production and distribution of resources that impacts consumers of content and larger cultural attitudes towards women. \ 


## Exploratory Data Analysis

Our data, as introduced above, includes information about popular movies. This is a tidy data set that has samples of 1,794 films released between the years of 1970 and 2013 @hickey2014dollars. To wrangle data’s variables, they were filtered to include “title”, “binary”, and  “budget”. The change to the buget_2013 column reflects the inflation adjusted budget calculated for all movies while title reveals the name of the movie and the binary column includes the pre-identified values of PASS/FAIL calculated with the Bechdel Test.\

Given the gender bias evident from the Bechdel Test results, our data analysis considers the relationship of the numeric variable of budget and the categorical variable of Pass/Fail to underrepresentation. We propose the question  "Do movies that pass the Bechdel Test have a higher budget?".  The numerical variable Budget serves as the outcome variable in our analysis while the categorical variable to Pass /Fail the Bechdel Test provides the explanatory variable of the analysis. To test the strength and magnitude of this relationship we use a regression model with a permutation test.

### Data Wrangling 

#### Imported Raw Data
```{r}
#| label: tbl-movies_raw
#| tbl-cap: "Raw Data Set Import "
#| eval: true
#| echo: false
#| message: false
movies_raw <- read_sheet("https://docs.google.com/spreadsheets/d/12xXTpjqwY35HV5WcAZtIGxnLc8xmBD0hQPVZKR5njcY/edit?gid=1554984819#gid=1554984819")
movies_raw_display <- slice(movies_raw, 1:3) 
kable(movies_raw_display) 
```
This @tbl-movies_raw displays the raw data we imported from @fivethirtyeight_bechdel to use for this project.

#### Data for Analysis
```{r}
#| label: tbl-selected_moviedata
#| tbl-cap: "Wrangled Movies Data for Research Project "
#| eval: true
#| echo: false
#| message: false
movies <- movies_raw |>
  select(title, binary, budget)

movies <- slice(movies, 1:5)
kable(movies)
```

```{r}
#| label: tbl-movies_clean_budget
#| tbl-cap: "Movies Clean Includes the 2013 inflation adjusted Budget"
#| eval: true
#| echo: false
#| message: false
movies_clean <- movies_raw |>
  select(c("title", "binary", "budget_2013")) |>
  mutate(budget_2013 = as.numeric(budget_2013))
movies_clean_display <- slice(movies_clean, 1:5)
kable(movies_clean_display)
```
Here @tbl-selected_moviedata we see the variables that we include for our analysis. \

In @tbl-movies_clean_budget, the Budget variable has been wrangled and updated to reflect the 2013 inflation adjusted budget.\

### Summary of Data
```{r}
#| label: tbl-movies_summary
#| tbl-cap: "Summary of Movies Data "
#| eval: true
#| echo: false
#| message: false
plotting <- movies_clean |>
  group_by(binary) |>
  summarize(avg_budget = mean(budget_2013),
           Number = n(),
           Standard_Deviation = sd(budget_2013))
kable(plotting)
```

```{r}
#| label: fig-movies_histogram
#| fig-cap: "Average Budget Differences"
#| eval: true
#| echo: false
#| message: false
ggplot(plotting, aes(binary, avg_budget)) +
  geom_bar(stat = "identity") +
  scale_y_continuous(labels = scales::comma)
```

```{r}
#| label: fig-movies_boxplot
#| fig-cap: "Boxplot showing the difference between average movie budgets more accuratly capturing outliers"
#| eval: true
#| echo: false
#| message: false
ggplot(movies_clean, aes(binary, budget_2013)) +
  geom_boxplot() +
  scale_y_continuous(labels = scales::comma)
```

In @tbl-movies_summary we see the breakdown of average budgets that either pass or fail the Bechdel test, the standard deviation is given along with the measure of Central Tendancy and the Number column reflects the number of movies in each category contained in the data set.\

Next @fig-movies_histogram we see the that this bar plot visualizes the differnce between average budgets. This origional finding prompted our hypotesis test to explore this difference between budgets and passing or failing the bechdel test.\

In @fig-movies_boxplot we can see the breakdown of this differnce in average budget more precisely with a boxplot that captures outliers.\

## Results

To exactly quantify the relationship between the average budget of a movie and whether it passes or fails the Bechdel test, we fit a categorical regression model. In the model we show in @tbl-regression_table, the outcome variable is the budget of a movie and the explanatory variable is the movie passing or failing the aforementioned test. The predictions created from the model predictions are visualized in @fig-Categorical_vizualization.


(do we need this code?)
### Hypothesis Testing
```{r}
#| label: fig-summary_stat
#| fig-cap: "Calculating the Summary Statistic for Hypothesis Test "
#| eval: true
#| echo: false
#| message: false
62911555 - 46274167
```

#### Categorical regression model
```{r}
#| label: tbl-regression_table
#| tbl-cap: "Regression Table"
#| eval: true
#| echo: false
#| message: false
bechdel_model <- lm(budget_2013 ~ binary, data = movies_clean)

bechdel_model_clean <- get_regression_table(bechdel_model)

kable(bechdel_model_clean)
```

#### Movie Budgets vs. Bechdel Test Binary
```{r}
#| label: fig-Categorical_vizualization
#| fig-cap: "Catagorical Visualization of difference between average movie budgets "
#| eval: true
#| echo: false
#| message: false
ggplot(bechdel_model, aes(x = binary, y = budget_2013)) +
  geom_jitter(width = 0.25) +
  geom_categorical_model() +
  labs(title = "Movie Budgets vs. Bechdel Test Binary",
       x = "Binary - PASS/FAIL",
       y = "Budget (in 2013 USD)") +
  theme_minimal() +
  scale_y_continuous(labels = scales::comma)
```

The @tbl-regression_table fits our hypothesis test to the data and gives us a table of values. The probability of observing a test statistic at least as extreme as the one observed from our sample, assuming the null hypothesis is true, is 0. Our alpha level is 0.05, and our p-value is 0, and because p-value \< alpha value, we can reject the null hypothesis and conclude that there is a statistical significance between the average budget and a movie passing or failing the bechdel test.\ 

One can see in @fig-Categorical_vizualization the Catagorical Visualization of difference between average movie budgets.\

#### Regression Equation:\
The regression equation created/described by the Table is:

$$
\widehat{Budget} = \widehat{\beta_0} + \widehat{\beta_1} \cdot 1_{Pass}(Binary)
$$

$$
\widehat{Budget} = 62911555 + (-16637388 * 1_{Pass}(Binary))
$$
The intercept, ($\widehat{\beta_0} = 62911555$) represents the average budget of a movie when it passes the bechdel test, and $\widehat{\beta_1}$, represents the decrease in the average budget when a movie fails the bechdel test, which is -16637388.


Although this might be due to chance, a permutation test on the model’s slope coefficient shows that there is not strong enough evidence to conclude that the differences between the average budgets are produced simply due to variability between the sample we used and others (t = 16637388, p = 0).


### Discussion

#### Conclusions

Through summarizing the data, we have gathered that, after adjusting for inflation in 2013 USD, the average budget of movies that fail the Bechdel test tends to be higher than the average budget of movies that pass it.

- talk about more insights with a more specific look at how other variables may be affected; do certain genres suffer from failing the test more than others? - talk about how it could be for many reasons that we, the researchers are still unsure about, like it could be that writers make deliberate decisions to give importance to main male leads rather than female leads, sexism within writing scripts, sexism from writers and/or executives that help make decisions on budgets. But it is a problem if films without big budgets behind them have actual 3-dimensional characters. Overall

#### Limitations

There are several limitations to the dataset and our findings that lead us to say that more research on the topic is needed to make any definitive conclusions. For starters, this dataset only uses movies from 1990 to 2013. There wasn’t a column in the dataset we used that defined where these movies were made and released as well, and while some international movies such as ____ and ___ are included, it may not give the best picture to analyze them in USD compared to their own country’s currency due to currency exchange rates and what’s considered a ‘high’ and ‘lower’ budget in their country’s of origins. This would be a good place to start with looking into this topic further, as it would give insight into whether or not more movies produced outside of the US fare better in terms of budget (in their country’s economy) compared to US film made and released in the US film market. With all this being said, it may not be appropriate to generalize these results to the global film industry as a whole, and movies made recently or before the 1990s. 

These results also don’t model how well the movie passes the test or provides insight to that; some movies that pass may consist of compelling and plot-relevant female characters that have conversations throughout the movie or simply have a single moment of them talking in an hour runtime about something irrelevant to the plot. The test is simple, and while a movie not passing it says a lot for the representation of women in that film (whether it exists or not), a movie passing doesn’t really say a lot about the quality of that representation. More qualitative research should be undertaken to gain a deeper insight. 

Additionally, it is important to note that this study does not take away from the fact that there are many factors that go into creating a budget, and we cannot claim that dimensionality of female characters is the sole or even main reason that a movie gets the budget that it does.

### Further Questions

On the last point, it should not understated the importance of doing further quantitative analysis into how other factors interact with a movie’s ability to pass or fail this test. For example, should a movie pass the bechdel test, does it get more or less gross domestic and/or international revenue? That could provide insight into how well certain movies are received by the public both within the US and outside of it.

        
One could also do a breakdown of the average budgets of movie vary between different ‘levels’ of passing the bechdel test. If there is no representation of women, if the women in the movie don’t talk, and if the movie passes all metrics of the bechdel test. Figure ___ shows an exploratory analysis of how the domestic gross varies between films that pass/fail the bechdel test, and figure  ___ is another exploratory analysis visualizing the possible link between different levels of passing/failing the bechdel test and a movie’s average budget. 


#### Bibliography









